A tutorial of Bayesian beta regressions with brms in R

### 2025-02-03 manuscript/manuscript.qmd
add some basic information on the beta distribution

---
title: A tutorial of Bayesian beta regressions with brms in R
author:
  - name: Stefano Coretta
    email: s.coretta@ed.ac.uk
    affiliations: 
        - id: uoe
          name: University of Edinburgh
          department: Linguistics and English Language
          address: 3 George Sq
          city: Edinburgh
          state: United Kingdom
          postal-code: EH8 9AD
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Paul Bürkner
    email: paul.buerkner@gmail.com
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: linguistics.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

@@

However, proportion are naturally not Gaussian, since they are limited between 0 and 1. The theoretical distribution that generates values with this characteristics is the beta distribution. Thus, regression models with proportions as the outcome variable should be fitted using a beta distribution as the distribution family. This tutorial introduces researchers to beta regression models in R using the package brms. Familiarity with regression modelling in R with a package like lme4 is assumed, but no prior knowledge of Bayesian statistics is necessary.

@@

...

>>>>

---
title: A tutorial of Bayesian beta regressions with brms in R
author:
  - name: Stefano Coretta
    email: s.coretta@ed.ac.uk
    affiliations: 
        - id: uoe
          name: University of Edinburgh
          department: Linguistics and English Language
          address: 3 George Sq
          city: Edinburgh
          state: United Kingdom
          postal-code: EH8 9AD
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Paul Bürkner
    email: paul.buerkner@gmail.com
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: 
  - linguistics.bib
  - statistics.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

@@

However, proportion are usually not Gaussian, since they are limited between 0 and 1. Thus, regression models with proportions as outcome variable should be fitted using a likelihood family that assumes such unit interval data. A common and flexible distribution for this purpose is the beta distribution. 
This tutorial introduces researchers to beta regression models in R using the package brms. Familiarity with regression modelling in R with a package like lme4 is assumed, but no prior knowledge of Bayesian statistics is necessary.

@@

The beta distribution is often considered as *the* distribution for the modeling of unit interval data [@ferrari_beta_2004; @cribari-neto_beta_2010]. It has been extensively studied theoretically [@krysicki_new_1999; @gupta_handbook_2014; @espinheira_beta_2008] and is used a baseline to compare other unit interval distributions against [@kieschnick_regression_2003; @bonat_regression_2013; @lopez_bayesian_2013]. Specifically in Bayesian regression models, it was shown to perform well across a wide range of scenarios [@scholz_posterior_2023; @scholz_prediction_2025], which is why we focus on the beta distribution in this tutorial.

We use a common mean parameterization for the beta distribution with mean parameter $\mu$, bounded between 0 and 1, and positive precision parameter $\phi$ that is roughly proportional to the inverse of the variance $\text{Var}(y) = \frac{\mu \, (1 - \mu)}{\phi + 1}$. That is, the larger $\phi$ the smaller the variance of the corresponding beta distribution. While the mathematical details are not needed to understand the content of this tutorial, we still write down the density below for reasons of completeness:

$$
p(y \mid \mu, \phi) = \frac{1}{B(\mu \, \phi, \; (1-\mu) \, \phi)} \; y^{\mu \, \phi -1} \; (1-y)^{(1-\mu) \, \phi -1},
$$

where $B$ is the Beta function, a complex integral for whose numerical approximation efficient algorithms
exist in every programming language.
